Active Learning from Blackbox to Timed Connectors

### 2016-01-16 paper/paper.tex
.

\subsection{Counter-Examples' Analysis} 
In our framework, the oracle function $eq$ is used to check the equivalence between hypothesis
models and blackboxes. It is also responsible for generating counter-examples.
Generally, as mentioned in \cite{DBLP:journals/iandc/Angluin87,DBLP:conf/sfm/SteffenHM11},
equivalence query has been proved impossible in blackbox models.

Apparently, the closed hypothesis presented in Figure \ref{fig:hypo2} is not equal to the 2-Timer
channel. It's easy to find a counter-example $s=a,\rnoread-T-T-\bot,\rread$ where $mq(s)=a$ while
according to the hypothesis, the result shoud have been $\rempty$. As mentioned above, we use a
suffix set $D$ to distinguish different states. The existance of counter-examples shows that the
current suffix set $D$ is not powerful enough to distinguish all different states. We need to analyze
the counter-example and in turn append more elements to $D$.

@@

We have programmed Reo channels as a new package in Golang. The package is well-written for not only
formal verification but also practical use.

@@

\subsection{Case Studies}
A simple example of timed connector is presented to show how L* works.
\begin{figure}[ht]
  \begin{center}
    \input{./images/expfifo.tex}
  \end{center}
  \caption{Expiring FIFO1 Channel (ExpFIFO1)}
  \label{fig:expfifo}
\end{figure}

@@

@@

\begin{table}[ht]
  \renewcommand{\arraystretch}{1.3}
  \caption{Time-Cost Analysis}
  \label{tabel:timecost}
  \centering
  \begin{tabular}{l||rrrr}
    \hline
    & FIFO & Alternator & Gate & ExpFIFO1\\
    \hline\hline
    Membership Query(s) & 41.571 & 126.468 & 169.161 &\\
    Hypothesis Query(s) & 0.001 & 0.003 & 0.004 &\\
    Total Time(s) & 41.715 & 165.114 & 247.098 &\\
    Membership Query(\%) & 99.6 & 76.6 & 68.5 &\\
    \hline
  \end{tabular}
\end{table}

@@

As mentioned above, our implementation in \texttt{Golang} is well-prepared not only for academic use
but also for practical concurrent programming. The following code shows how to compose an alternator
connector (see Figure \ref{fig:reoconnector}) in \texttt{Golang}. An intact version of this example
can be found in our github repo.

>>>>

\subsection{Counter-Examples' Analysis} 
Apparently, the closed hypothesis presented in Figure \ref{fig:hypo2} is not equal to the 2-Timer
channel. It's easy to find a counter-example $s=a,\rnoread-T-T-\bot,\rread$ where $mq(s)=a$ while
according to the hypothesis, the result shoud have been $\rempty$. In this section, we show how to
find and analyze counter-examples, which is mainly based on \cite{DBLP:conf/sfm/SteffenHM11}.

Firstly, we give a formal definition of \emph{counter-examples}. With a observation table $obs$ and
a sequence $s$, we use $acc(s)$ to denote the access sequence of $obs\xrightarrow{s}$ and $hq(s)$ to
denote the corresponding output.

\begin{definition}[Counter-Example]
  We use $hq$ to denote execution result of the hypothetical Mealy machine, sequence $s$ is called a
  counter-example iff $mq(s)\neq hq(s)$.
\end{definition}

Now we consider the reason that lead to existance of counter examples. Since
the suffix set $D$ is used to distinguish different states, the existance of counter-examples
shows that the current suffix set $D$ is not powerful enough to distinguish all different states.

Considering a counter example $s$, we have
\[
\mbox{MaxSuffix}(s)=\max_d\{s=s'd, mq(acc(s')d)\neq mq(s)\}
\]
Since $mq(acc(s))=hq(s)\neq mq(s)$ and $mq(s) = mq(s)$, the existance of such $d$ is easy to prove.
After appending $d$ in $D$, we explain the closedness will be violated.

The whole L* algorithm can be concluded as follows.
\begin{algorithm} 
  \caption{L*} 
  \label{alg:lstar}
  \KwIn{Oracle interface $mq,eq$, Input actions $\mathcal{I}$}
  \KwOut{Observation table $obs$} 
  $D=I$\;
  \Repeat{$ce=true$}{
    $obs$ = BuildTable($mq$, $\mathcal{I}$, $D$)\;
    $ce=eq(obs)$\;
    \If {$ce!=true$} {
      D.append(MaxSuffix(s))\;
    }
  }
  \Return $obs$\; 
\end{algorithm}

@@

@@

\subsection{Case Study}
A simple example of timed connector is presented to show how L* works.
\begin{figure}[ht]
  \begin{center}
    \input{./images/expfifo.tex}
  \end{center}
  \caption{Expiring FIFO1 Channel (ExpFIFO1)}
  \label{fig:expfifo}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \input{./images/expfifosemantics.tex}
  \end{center}
  \caption{Learn Result of The ExpFIFO1 where $n=2,\Sigma=\{a\}$}
  \label{fig:expfifosemantics}
\end{figure}

Figure \ref{fig:expfifosemantics} shows the learning result of this example.
To simplify the graph, we ignore the all the trivial transitions $\frac{\bot,\rnoread}{\rempty}$
and block transitions. More details of this case can be found in our \emph{github repo}.

@@

\begin{table}[ht]
  \renewcommand{\arraystretch}{1.3}
  \caption{Time-Cost Analysis}
  \label{tabel:timecost}
  \centering
  \begin{tabular}{l||rrr}
    \hline
    & FIFO & Alternator & Gate \\
    \hline\hline
    Membership Query(s) & 41.571 & 126.468 & 169.161 \\
    Hypothesis Query(s) & 0.001 & 0.003 & 0.004 \\
    Total Time(s) & 41.715 & 165.114 & 247.098 \\
    Membership Query(\%) & 99.6 & 76.6 & 68.5 \\
    \hline
  \end{tabular}
\end{table}

@@

Our implementation in \texttt{Golang} is well-prepared not only for academic use
but also for practical concurrent programming. The following code shows how to compose an alternator
connector (see in Figure \ref{fig:reoconnector}) in \texttt{Golang}. An intact version of this example
can be found in our github repo.
